---
title: "Linear Regression - Multivariate"
author: "Carlos Sanchez"
date: "19/02/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
n = 100
x <- rnorm(n)
x2 <- rnorm(n)
x3 <- rnorm(n)

y <- 1 + x + x2 + x3 + rnorm(n, sd = .1)
ey <- resid(lm(y ~ x2 + x3))
ex <- resid(lm(x ~ x2 + x3)) 
sum(ey * ex) / sum(ex^2)
coef(lm(ey ~ ex - 1))
coef(lm(y ~ x + x2 + x3))
```

#Example Mulivariate

```{r}
require(datasets)
data(swiss)
?swiss
require(GGally)   #Collection of add-ons tools for ggplot2 (pair, etc)
require(ggplot2)
library(dplyr)
```



```{r}
g <- ggpairs(swiss, lower = list(continuous = wrap("smooth", size = 0.1)), wrap = c(method = "loess"))
g
```



```{r}
summary(lm(Fertility ~ ., data = swiss))
summary(lm(Fertility ~ ., data = swiss))$coef
```



```{r}
summary(lm(Fertility ~ Agriculture, data = swiss))$coef


```

How to reverse the sign of a effect?

```{r}
n <- 100
x2 <- 1:n
x1 <- .01 * x2 + runif(n, -.1, .1)
y <-  -x1 + x2 + rnorm(n, sd = .01)

```



```{r}
summary(lm(y ~ x1))$coef
summary(lm(y ~ x1 + x2))$coef
```



```{r}
dat <- data.frame(y = y, x1 = x1, x2 = x2, ey = resid(lm(y ~ x2)), ex1 = resid(lm(x1 ~ x2)))

g <- ggplot(dat, aes(y = y, x = x1, color = x2)) +
  geom_point(color = "grey50", size = 2) +
  geom_smooth(method = lm, se = FALSE, color = "black") +
  geom_point(size = 2)
g
```



```{r}
g2 <- ggplot(dat, aes(y = ey, x = ex1, color = x2)) +
  geom_point(color = "grey50", size = 1.4) +
  geom_smooth(method = lm, se = FALSE, color = "black") +
  geom_point(size = 1.4)
g2

```


Problems of including innecesary variable (ex. one that is linear combination of other 2)
```{r}
z <-  swiss$Agriculture + swiss$Education
lm(Fertility ~ . + z, data = swiss)
```



```{r}
require(datasets)
data("InsectSprays")
require(stats)

```



```{r}
g <- ggplot(data = InsectSprays, aes(y = count, x = spray, fill = spray)) +
  geom_violin(color = "black", size = 0.5) +
  xlab("Type of spray") +
  ylab("Insect count")
g
```



```{r}
summary(lm(count ~ spray, data =InsectSprays))$coef
#14.5 is the mean of spray "A", rest Estimates are relative to A. p-value tells us if A from the rest are different
```


If we remove the intercept, then it includes A and the absolute values not relatives. In this case, p-value is compare to 0, instead of A, what doesn't give us much information

```{r}
summary(lm(count ~ spray - 1, data =InsectSprays))$coef
```

To use spray "C" as the reference level:

```{r}
spray2 <- relevel(InsectSprays$spray, "C")
summary(lm(count ~ spray2, data =InsectSprays))$coef

```

Create a binary variable

```{r}
swiss <- mutate(swiss, CatholicBin = 1 * (Catholic > 50))

```



```{r}
g <-  ggplot(swiss, aes(x = Agriculture, y = Fertility, colour = factor(CatholicBin))) +
  geom_point(size = 2, color = "black") +
  geom_point(size =1.5) +
  xlab("% in Agriculture") +
  ylab("Fertility")
g
```


Case when we consider interaction between variables 

1. Model without taking into account type of religion

```{r}
fit <- lm(Fertility ~ Agriculture, swiss)
g1 <- g +
  geom_abline(intercept = coef(fit)[1], slope = coef(fit)[2], size = 1.3)
g1
```



```{r}
summary(fit)$coef
```

2. Model considering type of religion but no interaction (parallel slope)

```{r}
fit <- lm(Fertility ~ Agriculture + factor(CatholicBin), swiss)
summary(fit)$coef

```



```{r}
g1 <- g +
  geom_abline(intercept = coef(fit)[1], slope = coef(fit)[2], size = 1.3) +
  geom_abline(intercept = coef(fit)[1] + coef(fit)[3], slope = coef(fit)[2], size = 1.3)

g1
```


3. Model considering type of religion but WITH interaction (different slope)

```{r}
fit <- lm(Fertility ~ Agriculture * factor(CatholicBin), swiss)  # change the '+' by '*'
summary(fit)$coef
```

```{r}
g1 <- g +
  geom_abline(intercept = coef(fit)[1], slope = coef(fit)[2], size = 1.3, color= "red") +
  geom_abline(intercept = coef(fit)[1] + coef(fit)[3], 
              slope = coef(fit)[2] + coef(fit)[2], 
              size = 1.3, color ="blue")

g1
```